**NAME OF THIS STUDY**

### 2018-11-08 submission/manuscript.Rmd
comments added

```{r knitr_settings, eval=TRUE, echo=FALSE, cache=FALSE,  warning=FALSE, message=FALSE}
# Read in necessary libraries
deps = c("doParallel","pROC", "caret","knitr","rmarkdown","vegan","gtools", "tidyverse");
for (dep in deps){
  if (dep %in% installed.packages()[,"Package"] == FALSE){
    install.packages(as.character(dep), quiet=TRUE);
  }
  library(dep, verbose=FALSE, character.only=TRUE)
}
opts_chunk$set("tidy" = TRUE)
opts_chunk$set("echo" = FALSE)
opts_chunk$set("eval" = TRUE)
opts_chunk$set("warning" = FALSE)
opts_chunk$set("cache" = FALSE)

@@

	if(is.numeric(x)){
		if(abs(x - round(x)) < .Machine$double.eps^0.5){
			paste(format(x,big.mark=',', digits=0, scientific=FALSE))
		} else {
			paste(format(x,big.mark=',', digits=1, nsmall=1, scientific=FALSE))
		}
	} else {
    	paste(x)      
	}
}
knitr::knit_hooks$set(inline=inline_hook)
options(kableExtra.latex.load_packages = FALSE)
options(kableExtra.auto_format = FALSE)
options(knitr.table.format = "latex")
library(kableExtra)
```

```{r LoadData, eval=TRUE, echo=FALSE, cache=FALSE, warning=FALSE, message=FALSE}
# Read in metadata and select only sample Id and diagnosis columns
meta <- read.delim('../data/metadata.tsv', header=T, sep='\t') %>%
  select(sample, dx)

@@

# Read in AUCs table generated from l2 logistic regression model for all samples
  # Carcinomas + Adenomas are 1 and Normal is 0 for binary predictive model
  # FIT is a feature
  # The scaler is MinMax scaler
logit <- read.delim('../data/process/L2_Logistic_Regression.tsv', header=T, sep='\t') %>%
  select(level_1, AUC) %>% 
  rename(Performance = level_1) %>% 
  mutate(model="L2-Logistic Regression")

# Read in AUCs table generated from l1 SVM linear kernel for all samples
  # Carcinomas + Adenomas are 1 and Normal is 0 for binary predictive model
  # FIT is a feature
  # The scaler is StandardScaler
l1svm <- read.delim('../data/process/L1_SVM_Linear_Kernel.tsv', header=T, sep='\t') %>%
  select(level_1, AUC) %>% 
  rename(Performance = level_1) %>% 
  mutate(model="L1-SVM Linear")

# Read in AUCs table generated from l2 SVM linear kernel for all samples
  # Carcinomas + Adenomas are 1 and Normal is 0 for binary predictive model
  # FIT is a feature
  # The scaler is StandardScaler
l2svm <- read.delim('../data/process/L2_SVM_Linear_Kernel.tsv', header=T, sep='\t') %>%
  select(level_1, AUC) %>% 
  rename(Performance = level_1) %>% 
  mutate(model="L2-SVM Linear")

# Read in AUCs table generated from  SVM RBF kernel for all samples
  # Carcinomas + Adenomas are 1 and Normal is 0 for binary predictive model
  # FIT is a feature
  # The scaler is StandardScaler
svmRBF <- read.delim('../data/process/SVM_RBF.tsv', header=T, sep='\t') %>%
  select(level_1, AUC) %>% 
  rename(Performance = level_1) %>% 
  mutate(model="SVM RBF")

# Read in AUCs table generated from xgboost for all samples
  # Carcinomas + Adenomas are 1 and Normal is 0 for binary predictive model
  # FIT is a feature
  # The scaler is MinMaxScaler
xgboost <- read.delim('../data/process/XGBoost.tsv', header=T, sep='\t') %>%
  select(level_1, AUC) %>% 
  rename(Performance = level_1) %>% 
  mutate(model="XGBoost")

# Read in AUCs table generated from random forest for all samples
  # Carcinomas + Adenomas are 1 and Normal is 0 for binary predictive model
  # FIT is a feature
  # The scaler is MinMaxScaler
rf <- read.delim('../data/process/Random_Forest.tsv', header=T, sep='\t') %>%
  select(level_1, AUC) %>% 
  rename(Performance = level_1) %>% 
  mutate(model="Random Forest ")

# Read in AUCs table generated from decision tree for all samples
  # Carcinomas + Adenomas are 1 and Normal is 0 for binary predictive model
  # FIT is a feature
  # The scaler is MinMaxScaler
dt <- read.delim('../data/process/Decision_Tree.tsv', header=T, sep='\t') %>%
  select(level_1, AUC) %>% 
  rename(Performance = level_1) %>% 
  mutate(model="Decision Tree")

# Put  all the AUCs for all the models together and group by model
all <- bind_rows(logit, l1svm, l2svm, svmRBF, xgboost, rf, dt) %>% 
  group_by(model)
```

```{r table_1_models, echo=FALSE, results='asis'}
# We load a model parameters file to make our Table 1. Model parameters file has the information of optimized hyper-parameters, pre-processing and cross-validation methods and software implementation for all the classification algorithms.

# This file is hand-made. Therefore needs to be changed manually if new information wants to be added or information needs to be changed.

model_param <- read.table('../data/process/model_parameters.txt', header=T, sep='\t') %>% 
   select(Method, Parameter, Cross.Validation, Epochs, Scaler, Sklearn.Function) %>% 
  rename("Cross Validation"="Cross.Validation") %>%
  rename("Epoch"="Epochs") %>%
  rename("Sklearn Function"="Sklearn.Function")

             ### Table 1 generation ###

@@

```{r table_2_3_params, echo=FALSE, results='asis'}
# We load a param_grid.scv file to make our Table 2 and 3. This file has the information of the ranges of the optimized hyper-parameters for each model.

@@

@@

              ## Table 2 generation ##
Table2 <- kable(linear_param_range, "latex", caption = "The range of optimized hyper-parameters for logistic regression and support vector machines.", booktabs = T, col.names = NULL, linesep = "") %>%
  # latex options make table to be condensed
  # table itseld if left aligned
  kable_styling(latex_options = c("striped", "scale_down"), position = "left") %>% 
  # the column names are removed when the file is read so I put headers above the grouped/replicated modeling method for
  add_header_above(c("Parameter", "L2 Logistic" = 4, "L1 SVM Linear" = 4, "L2 SVM Linear " = 3, "SVM RBF"=5), bold=TRUE) %>% 
  # Put a left border to seperate above header lines
  column_spec(2, border_left = T) %>%
  column_spec(6, border_left = T) %>% 
  column_spec(10, border_left = T) %>% 
  column_spec(13, border_left = T)

@@

                  ## Table 3 generation ##
Table3 <- kable(tree_param_range, "latex", caption = "The range of optimized hyper-parameters for tree based classification algorithms.", booktabs = T, col.names = NULL, linesep = "") %>%
kable_styling(latex_options = c("striped", "scale_down"), position = "left", font_size=6) %>% 
  add_header_above(c("Parameter", "Random Forest"=5, "Decision Tree"= 4, "XGBoost"=3 ), bold=TRUE) %>% 
  column_spec(2, border_left = T) %>% 
  column_spec(7, border_left = T)%>% 
  column_spec(11, border_left = T)
```
\vspace{35mm}

>>>>

```{r knitr_settings, eval=TRUE, echo=FALSE, cache=FALSE,  warning=FALSE, message=FALSE}
######################################################################
#----------------- Read in necessary libraries -------------------#
######################################################################
deps = c("doParallel","pROC", "caret","knitr","rmarkdown","vegan","gtools", "tidyverse");
for (dep in deps){
  if (dep %in% installed.packages()[,"Package"] == FALSE){
    install.packages(as.character(dep), quiet=TRUE);
  }
  library(dep, verbose=FALSE, character.only=TRUE)
}
######################################################################
#-------------- Define the chunk options ----------------#
######################################################################
opts_chunk$set("tidy" = TRUE)
opts_chunk$set("echo" = FALSE)
opts_chunk$set("eval" = TRUE)
opts_chunk$set("warning" = FALSE)
opts_chunk$set("cache" = FALSE)

@@

	if(is.numeric(x)){
		if(abs(x - round(x)) < .Machine$double.eps^0.5){
			paste(format(x,big.mark=',', digits=0, scientific=FALSE))
		} else {
			paste(format(x,big.mark=',', digits=1, nsmall=1, scientific=FALSE))
		}
	} else {
    	paste(x)      
	}
}
######################################################################
#-------------- Define knitr options to allow kable ----------------#
######################################################################
knitr::knit_hooks$set(inline=inline_hook)
options(kableExtra.latex.load_packages = FALSE)
options(kableExtra.auto_format = FALSE)
options(knitr.table.format = "latex")
library(kableExtra)
```

```{r LoadData, eval=TRUE, echo=FALSE, cache=FALSE, warning=FALSE, message=FALSE}
######################################################################
#----------------- Load OTU table and MetaData -----------------#
######################################################################

# Read in metadata and select only sample Id and diagnosis columns
meta <- read.delim('../data/metadata.tsv', header=T, sep='\t') %>%
  select(sample, dx)

@@

######################################################################
#------------ Load .tsv data generated in Python -----------------#
######################################################################

# Read in AUCs table generated from L2 logistic regression model for all samples:
#       Carcinomas + Adenomas are 1 and Normal is 0 for binary.
#       FIT is a feature
#       The scaler is MinMax scaler
logit <- read.delim('../data/process/L2_Logistic_Regression.tsv', header=T, sep='\t') %>%
  select(level_1, AUC) %>% 
  rename(Performance = level_1) %>% 
  mutate(model="L2-Logistic Regression")

# Read in AUCs table generated from l1 SVM linear kernel for all samples
#       Carcinomas + Adenomas are 1 and Normal is 0 for binary.
#       FIT is a feature
#       The scaler is Standard scaler
l1svm <- read.delim('../data/process/L1_SVM_Linear_Kernel.tsv', header=T, sep='\t') %>%
  select(level_1, AUC) %>% 
  rename(Performance = level_1) %>% 
  mutate(model="L1-SVM Linear")

# Read in AUCs table generated from l2 SVM linear kernel for all samples
#       Carcinomas + Adenomas are 1 and Normal is 0 for binary.
#       FIT is a feature
#       The scaler is Standard scaler
l2svm <- read.delim('../data/process/L2_SVM_Linear_Kernel.tsv', header=T, sep='\t') %>%
  select(level_1, AUC) %>% 
  rename(Performance = level_1) %>% 
  mutate(model="L2-SVM Linear")

# Read in AUCs table generated from  SVM RBF kernel for all samples
#       Carcinomas + Adenomas are 1 and Normal is 0 for binary.
#       FIT is a feature
#       The scaler is Standard scaler
svmRBF <- read.delim('../data/process/SVM_RBF.tsv', header=T, sep='\t') %>%
  select(level_1, AUC) %>% 
  rename(Performance = level_1) %>% 
  mutate(model="SVM RBF")

# Read in AUCs table generated from xgboost for all samples
#       Carcinomas + Adenomas are 1 and Normal is 0 for binary.
#       FIT is a feature
#       The scaler is MinMax (0-1) scaler
xgboost <- read.delim('../data/process/XGBoost.tsv', header=T, sep='\t') %>%
  select(level_1, AUC) %>% 
  rename(Performance = level_1) %>% 
  mutate(model="XGBoost")

# Read in AUCs table generated from random forest for all samples
#       Carcinomas + Adenomas are 1 and Normal is 0 for binary.
#       FIT is a feature
#       The scaler is MinMax (0-1) scaler
rf <- read.delim('../data/process/Random_Forest.tsv', header=T, sep='\t') %>%
  select(level_1, AUC) %>% 
  rename(Performance = level_1) %>% 
  mutate(model="Random Forest ")

# Read in AUCs table generated from decision tree for all samples
#       Carcinomas + Adenomas are 1 and Normal is 0 for binary.
#       FIT is a feature
#       The scaler is MinMax (0-1) scaler
dt <- read.delim('../data/process/Decision_Tree.tsv', header=T, sep='\t') %>%
  select(level_1, AUC) %>% 
  rename(Performance = level_1) %>% 
  mutate(model="Decision Tree")

######################################################################
#------------ Put all the loaded AUC tables together -----------------#
######################################################################
all <- bind_rows(logit, l1svm, l2svm, svmRBF, xgboost, rf, dt) %>% 
  group_by(model)
```

```{r table_1_models, echo=FALSE, results='asis'}
######################################################################
# We load a model parameters file to make our Table 1. Model parameters file has the information of optimized hyper-parameters, pre-processing and cross-validation methods and software implementation for all the classification algorithms.

# This file is hand-made. Therefore needs to be changed manually if new information wants to be added or information needs to be changed. 
######################################################################
model_param <- read.table('../data/process/model_parameters.txt', header=T, sep='\t') %>% 
   select(Method, Parameter, Cross.Validation, Epochs, Scaler, Sklearn.Function) %>% 
  rename("Cross Validation"="Cross.Validation") %>%
  rename("Epoch"="Epochs") %>%
  rename("Sklearn Function"="Sklearn.Function")

#------------------------Table 1 generation------------------------#

@@

```{r table_2_3_params, echo=FALSE, results='asis'}
######################################################################
# We load a param_grid.scv file to make our Table 2 and 3. This file has the information of the ranges of the optimized hyper-parameters for each model.

@@

# 1) First we will generate a Table for non-tree based models that shows C(penalty) and gamma hyperparameters.

# 2) Second we will generate a Table for tree based models that show all the hyperparameters used for it. 
######################################################################

@@

#------------------------Table 2 generation------------------------#
Table2 <- kable(linear_param_range, "latex", caption = "The range of optimized hyper-parameters for logistic regression and support vector machines.", booktabs = T, col.names = NULL, linesep = "") %>%
  # latex options make table to be condensed
  # table itseld if left aligned
  kable_styling(latex_options = c("striped", "scale_down"), position = "left") %>% 
  # the column names are removed when the file is read so I put headers above the grouped/replicated modeling method for
  add_header_above(c("Parameter", "L2 Logistic" = 4, "L1 SVM Linear" = 4, "L2 SVM Linear " = 3, "SVM RBF"=5), bold=TRUE) %>% 
  # Put a left border to seperate above header lines
  column_spec(2, border_left = T) %>%
  column_spec(6, border_left = T) %>% 
  column_spec(10, border_left = T) %>% 
  column_spec(13, border_left = T)

@@

#------------------------Table 3 generation------------------------#
Table3 <- kable(tree_param_range, "latex", caption = "The range of optimized hyper-parameters for tree based classification algorithms.", booktabs = T, col.names = NULL, linesep = "") %>%
kable_styling(latex_options = c("striped", "scale_down"), position = "left", font_size=6) %>% 
  add_header_above(c("Parameter", "Random Forest"=5, "Decision Tree"= 4, "XGBoost"=3 ), bold=TRUE) %>% 
  column_spec(2, border_left = T) %>% 
  column_spec(7, border_left = T)%>% 
  column_spec(11, border_left = T)
```
\vspace{35mm}
